Fine-scale Twitter Big Data Reveals Neighborhood Inequalities in Mental Health Effects of High Temperatures

### 2021-05-12 doc/temp-sentiment.tex
Updates from Overleaf

\begin{abstract}
%Limit 150 words: https://www.nature.com/nclimate/about/content
%Actually there are lots of examples of articles with > 150 words, but I think we should keep it concise
Higher temperatures associated with climate change are expected to have major impacts on human mental health.  Indeed, traditional analyses of heat and mental health outcomes using data collected by public agencies have found associations between elevated temperatures and increases in outcomes such as suicides and mental-health related hospitalizations.  However, these studies, which typically use data reported at the city or county level, have found no difference in vulnerability based on income.  We overturn this finding and show that there are stark differences in vulnerability using a novel data source: expressed sentiment in 250 million geolocated tweets, matched with prevailing weather as well as neighborhood economic and demographic conditions.  We find that increased temperatures worsen expressed sentiment in all areas, but that this effects is much stronger in poor and black neighborhoods.

@@

Keywords: Text Data, Environmental Shocks, Social Media Mining, Sentiment Analysis, Topic Modeling, Social Media, Public Opinion, Perception, Sentiment, Media, News, Twitter, Climate Change, Science Communication, Policy, Planning, Methods, Text Mining, Natural Language Processing, NLP, Latent Dirichlet Allocation, LDA, Political Ecology, Politics, Resilience, Adaptation, Suicide, Mortality, Morbidity, Self-harm,

%Cite this new paper: https://doi.org/10.1016/S2542-5196(20)30251-5
% And: https://eos.org/articles/long-term-drought-harms-mental-health-in-rural-communities and this is the conf. paper citation https://agu.confex.com/agu/fm20/meetingapp.cgi/Paper/771750
\section{Introduction}

% Organize as follows:

% Heatwaves are known to worsen mental health
% AND mental health is strongly associated with income
% AND wealthier better able to cope with heatwaves 
% BUT we know little about the interaction between heatwaves, mental health and income
% THEREFORE: we did this really novel study.

% Jeremiah Osborne-Gowey: Randy Olsen (of "Don't be such a scientist" fame) and his ABT - And But Therefor framework for communicating - https://www.huffpost.com/entry/and-but-therefore-randy-o_b_8813330
% Jeremiah Osborne-Gowey: https://blogs.scientificamerican.com/observations/how-the-word-but-could-save-the-world/

This work builds on existing work by \citep{baylis_weather_2018}.

%Intro Paragraph
% I think this should be more about climate change and mental health broadly, as well as a bit about how income has been theorized to affect vulnerability, but there is no evidence yet.  This can be shorter as well, as we have a lot of detail in the Background section 
% Cite this high-level study: https://www.nature.com/articles/s41558-018-0102-4/
% Also this review article saying more research is needed https://www.sciencedirect.com/science/article/pii/S0033350618302130
% -Matt

Research has shown strong linkages between ambient environmental conditions and mental health, with extreme temperatures  frequently associated with poorer overall mental health. Individual's mental health is strongly associated with income with people in lower income levels frequently experiencing higher stress levels and less able to find help for coping with these stresses. Other research on how people cope with environmental temperature extremes indicate that wealthier are better able to cope with heatwaves, whether from traveling to cooler climes, accessing air conditioning, or other temperature ameliorating strategies that may require access to financial capital. Yet we know relatively little about the interaction between temperature, mental health and income. Furthermore, previous studies that did attempt to examine relationships between environmental extremes and health sometimes aggregated data at scales too coarse to examine whether income may be an ameliorating factor. This study examines the interactions between these three using relatively fine-scale, publicly available data. Our research expands on existing studies of temperature and sentiment - one measure of mental health - by examining finer-scale data and incorporating income data to examine correlations between temperature extremes and sentiment.

Human moods and mental states are important aspects of overall well-being and can be influenced by ambient, persistent and fluctuating environmental conditions. Climate change is accelerating the rate and variability around meteorological norms like the timing, magnitude, intensity and duration of precipitation events and air temperature minima and maxima \href{https://www.ipcc.ch/site/assets/uploads/2018/03/SREX-Chap3_FINAL-1.pdf}{(link)}. Previous work \citep{baylis_weather_2018} indicate important linkages between human mood and meteorological conditions. Previous analyses of the effects of weather on mood (as expressed in sentiment of text-based message), however, were constructed at the aggregated city/day level and did not account for differences in socio-economic status which may affect access to resources (e.g., air conditioning) that can offset the effects of exposure to temperature extremes. Here, we build on these previous studies by examining the hourly effects of weather on human sentiment, an expression of mood, exploring geographic and economic heterogeneity at a finer-scale resolution than previous studies.

% https://medinform.jmir.org/2020/1/e16023/?utm_source=TrendMD&utm_medium=cpc&utm_campaign=JMIR_TrendMD_0
\section{Background}
\subsection{Temperature and Mental Health}
%Jeremiah Osborne-Gowey: We report on physiological and mental health in text here. Maybe just a heading of "health"?
%I think physical health is beyond the scope of this paper -Matt

%https://www.pnas.org/content/115/43/10953
Higher temperatures are associated with many indicators of worsened mental health.  Multiple studies have found strong evidence that higher temperatures are associated with increases in suicides in the United States \citep{Burke2018Aug, Mullins2019Dec, Dixon2007May}, and many others have demonstrated the same relationship around the world \citep{Qi2014Dec, Page2007Aug, Likhvar2011Jan}.  In addition to suicide, other research has found effects temperature on hospitalization events related to mental health disorders such as bipolar disorder and schizophrenia \citep{Mullins2019Dec, Lee2007Jan, Shapira2004Feb, Sung2013Feb, Gupta1992Jun, Hansen2008Oct}.

Temperature has been hypothesized to impact mental health through a number of pathways.  Work on biological mechanisms emphasize that there may be mental health effects from maintaining stable body temperature in high heat \cite{Lohmus2018Jul}.  Additionally, recent studies have found that nighttime temperatures affects sleep quality \citep{Obradovich2017May, Mullins2019Dec}

% The pathways
Higher temperature associated with climate change could influence mental health directly by exposing people to trauma. It could also indirectly influence mental health by affecting (1) physical health and (2)community well-being \citep{RN1314}. For the direct influence pathway, extreme heat events or increasing temperatures have been associated with the increase in aggression,higher suicide rates and other hospital admissions(Anderson and Anderson 1998,Cheatwood 1995; Cohn et al. 2004,Bouchama et al.2007.). Heat exposures in working environments could also reduce people's capacity to deal with physical and mental task, and increase the risk of accidents, because of the excessive core body temperature and dehydration((Bridger 2003; Kerslake 1972; Kjellstrom et al.2009d,Leithead and Lind 1964,Schrier et al. 1970). The loss of work capacity would in turn result in loss of income, which could also cause mental health problems(Kjellstrom 2009b).%% more physiological terms

As for the indirect pathway, first, because of the reciprocal causation relationship between physical health problems and mental health problems(Miller et al. 2009; Prince et al. 2007), heat exposures associated with climate change along with other climate events and indirect health risks threatening physical health will directly influence mental health((Berry 2007,World Health Organization 2008b). Second，disordered temperature associated with climate change may destroy the economic in agricultural-production-dependent communities. For example, extreme heat reduces the work capacity of laborers in farm fields(Kjellstrom et al. 2009d), which further destroys agricultural-supported industries in local area(Berry et al. 2008). The following economic pressures would undermine social capital and then lead to mental health problems.

% Heatwaves are known to worsen mental health

https://doi.org/10.1016/j.puhe.2018.06.008 (review study - will have more to cite)
https://www.nature.com/articles/s41558-018-0222-x
https://ehp.niehs.nih.gov/doi/full/10.1289/ehp.11339
https://ijmhs.biomedcentral.com/articles/10.1186/s13033-018-0210-6
https://www.nature.com/articles/s41558-018-0102-4/
https://doi.org/10.1016/j.jhealeco.2019.102240

[ADD SUICIDE DATA TO HEALTH SECTION?]
https://www.psychologytoday.com/us/blog/greening-the-media/201809/global-warming-and-suicide

\subsection{Mental Health and Vulnerability}
% This is good, but probably should be shortened.  Maybe one para showing that income matters, one para discussing pathways, also maybe focus/change phrasing to be more on "vulnerability" -Matt

% AND mental health is strongly associated with income
% 1）Childhood 2）adulthood 3）gender
% Perceived income level

% Talk more about race
% This is interesting: https://www.pnas.org/content/118/17/e2019624118.abstract?etoc

The relationship between income, socioeconomic status (SES), well-being and mental health is one of the strongest established patterns in psycho-social literature \citep{Easterlin1974Jan, holzer1986increased, Perry1996Sep} (I cant find Ng et al. 2014). Much of the research into these patterns focuses on the inequalities and distribution of the effects on individual well-being. Subsequent work on mental health outcomes, one measure of well-being, have established strong associations between SES, income and mental health with the most and least privileged most commonly associated with the best and worse mental health experiences and outcomes, respectively \citep{Sevenson2008Aug}(cites: Ng et al. 2014).

For example, results from a cross-sectional comparative study of socioeconomic factors and use of mental health services by people living in Ontario, Canada and the United States, indicate clear disparities in use of mental health services between those with the highest income levels and lowest mental health morbidity and those with the lowest income levels and highest mental health morbidity - those with higher incomes tended to have lower mental health morbidity issues while utilizing mental health services more than those with lower incomes and higher mental health issues \citep{Katz1997}.

Results from a systematic review of the literature on associations between social inequalities and mental health disorders indicate a clear and prevailing link between  one or more indicators of less social privilege and higher prevalence of mental health disorders \citep{Fryers2003}, with low income one of the most consistent markers of and associations with increases in common mental health disorders. Collectively, results from the income and mental health literature indicate that socially disadvantaged populations experience significantly more frequent common mental health disorders. Scholars use mental health disparities to indicate the disproportionate amount of mental health disorders among persons of low SES \citep{RN1292}

The established association between poverty and mental health revealed that that income distribution may have a significant influence upon mental health over and above the effect of poverty \citep{HANANDITA201459}.Poverty has been identified as one of the major risk factors in mental health. For instance, some systematic reviews have shown that socioeconomically disadvantaged children and adolescents were 2-3 times more likely to have mental health problems than others\citep{REISS201324}, while the household increases in financial resources are generally associated with an overall reduction of mental health problems\citep{2015Does}. Others have argued that increases in SES can serve as a buffer against the negative impacts of difficult life experiences on mental health, particularly for individuals already under substantive mental stress \citep{Kawachi2001Sep}. Thus, SES and income are associated with mental health improvements in SES and income can serve to decrease mental health problems.

Several reasons could explain why poverty may effect mental health. 1) the “social causation hypothesis” which suggests that stress or deprivation or decreasing the likelihood of people getting treatment may lead to poor mental health\citep{mills2015}, while social capital could reduce the likelihood of living in poverty and reduces the risk of mental disorders such as depressive symptoms and suicidal tendencies\citep{RN1291}. 2) negative life events in a person’s life such as job loss are also associated with the risks of poverty\citep{RN1293} and mental health\citep{TAMPUBOLON201420}, and the causal inferences on the effect of poverty and mental health had also been made in several studies. For example, Tampubolon and Hanandita\citep{TAMPUBOLON201420} find that individual social capital is positively associated with mental health while adverse events were negatively associated; Chang.et.al \citep{RN1291} find that subjective and objective poverty is significantly associated with a higher risk of adverse life events, less social support and mental distress. Negative life events and social support in serial mediate the relationship between subjective poverty and mental health.

% jieliu.cnah: Could delete
For children, poverty and poor maternal mental health often co-exist and are two of the risk factors for child development\citep{LUND20111502}, and maternal mental health is significantly associated with child general psychopathology\citep{ RN1289}. Researchers also find that transition into poverty increased children’s socioemotional behavior problems and maternal psychological distress \citep{WICKHAM2017e141}.

In general, the relationship between mental health and income level is well established. In the study, we explore whether income level would influence the link between weather and the sentiment in Twitter-posed text.

https://jamanetwork.com/journals/jamapsychiatry/fullarticle/211213
https://doi.org/10.1002/9781118410868.wbehibs570
https://www.sciencedirect.com/science/article/abs/pii/S027795362030527X
https://www.sciencedirect.com/science/article/pii/S2468266717300117
https://doi.org/10.1016/S2468-2667(17)30011-7
https://link.springer.com/article/10.1007/s00127-017-1370-4
 % Jeremiah Osborne-Gowey: Not as relevant. Saved the citation but didn't find enough relevant text to include reference to here in our manuscript. Maybe this: https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0116820
https://www.sciencedirect.com/science/article/abs/pii/S0143622814002537
also this: https://www.sciencedirect.com/science/article/abs/pii/S0140673614614604

\subsection{Income and Temperature}
% AND wealthier better able to cope with heatwaves 
https://www.nature.com/articles/nclimate3253
https://www.sciencedirect.com/science/article/abs/pii/S0378778818321327
https://www.mdpi.com/2225-1154/8/1/12/htm

NOT CLEAR: https://doi.org/10.1016/j.jhealeco.2019.102240 and https://www.nature.com/articles/s41558-018-0222-x find little evidence of adaptation

% BUT we know little about the interaction between heatwaves, mental health and income

% THEREFORE: we did this really novel study.
https://www.nature.com/articles/s41598-017-12961-9

%OTHER LEFTOVER TEXT TO MAYBE BRING IN LATER

Meteorological conditions can impact human physical (cite) and emotional states (cite). Emotional states and well-being are associated with physiological functioning and mental acuity which can affect social relationships, workplace productivity (cite) and health risks (cite). People that are more reliant on 1) livelihoods which require them to be outdoors (e.g., farming, agriculture, logging, etc.), 2) living and working in places where they are exposed to environmental minima and maxima, or 3) with prolonged exposure to ambient meteorological extremes are particularly vulnerable to changes in environmental exposure \citep{frimpong_heat_2017} and associated health risks (cite). Exposure to extreme environmental conditions can also have implications for workplace productivity and livelihoods \cite{kjellstrom_impact_2016}. For example, Nigerian maize farmers experience significant declines in productivity (2-8percent per degree above 17degC) along with substantive increases in health impacts including dehydration, muscle cramps, headaches and dizziness, heat exhaustion, sun stroke, and even death \cite{sadiq_impact_2019}. Other research indicates farm health and labor productivity were compromised under extreme heat and cold events in the Nepali Food Bowl region \cite{budhathoki_socio-economic_2019}. Similarly, Oregon farmers reported health impacts of working in both hot and cold conditions but the negative impacts were attenuated relative to the high heat situations \cite{bethel_heat-related_2014}. Other research found occupational injuries in Thailand increased with increasing heat exposure and occupational heat stress \cite{tawatsupa_association_2013}. Heat exposure risks are also a key factor in urban areas with particular attention needed for how the impacts of changing climate play out in health inequality \cite{friel_urban_2011}.

Current estimates place annual workplace productivity losses due to heat exposure at 15-20percent with that rate potentially doubling by 2050 \cite{kjellstrom_heat_2016}. Projected increases in future ambient temperature suggest serious short- and long-term potential health consequences from exposure to heat stress (cite) with some estimates putting potential productivity losses at several percentage points by 2030 \cite{kjellstrom_heat_2016}, with middle- and low-income countries particularly vulnerable as they are often more reliant on physical work for their livelihoods.

The level of changes in weather influences people’s mental health and patterns of emotions expressed. A study by Sun et al. 2018 \href{file:///Users/portia/Downloads/ijerph-16-00086-v2 20(1).pdf.}{(Portia has this PDF?)} demonstrates varied relationships between haze and negative emotions of the public under different seasons of the year. Being sad or happy influences one's style and value of reasoning. Time of day, season, location, and climate allow aggregate prediction of sentiments \cite{hannak_tweetin_2012} \href{https://www.ccs.neu.edu/~amislove/publications/Weather-ICWSM.pdf}{(link to PDF)} while positive affect from an evaluative statement enhances potential responses \cite{clore_how_2007} \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2483304/pdf/nihms40349.pdf}{(link to PDF)}. For instance, changes in temperature affects people’s response about weather. Depending on a reference temperature for an  area and time of year, people are more likely to comment on unusual weather for a particular place and time than on the same weather considered typical in another place \cite{moore_rapidly_2019}.

% Jeremiah Osborne-Gowey: Remove parts of this paragraph that relate to call for international research on these linkages as we now are only focusing on USA for Tweets.
Although climate change is an international phenomenon experienced in every country, its discussions vary from country to country. According to Vu and others \cite{vu_nationalizing_2019}, a country’s climate severity, economic status and governance determine variations in its media discussions. 
% Jeremiah Osborne-Gowey: I think we can remove this entire paragraph now as it's outside the focus of this paper. Might be relevant for our 2nd paper!
When removing the paragraph, cut from here and copy to the end of the doc where we have extra text saved. [DO NOT DELETE, just move from here]

% Jeremiah Osborne-Gowey: This is perhaps the only piece of this paragraph still relevant for this paper...
For example, Park and others \cite{park_mood_2013} \href{https://pdfs.semanticscholar.org/b282/feb759e57530b115dcc4bb080f96598a5246.pdf}{(link)} demonstrated that many people living in a state with higher mean temperature express more positive emotions on Twitter than those living in colder states in the US. Schmidt and others \cite{schmidt_media_2013} posit that climate severity and environmental factors related to carbon dependency influences the amount of coverage climate change receives in the media in countries like USA, Australia and Germany. Studies comparing how different countries portray climate change have been widely conducted in USA, UK, France and the Netherlands \cite{vu_nationalizing_2019}. Such comparisons are mainly between developed countries with developing countries outside the focus.  Schafer and O’Neil \cite{schafer_what_2013} advance the need for academic scholars to investigate transnational contexts within climate communication. Wealthier countries with comparatively higher GDP are likely to frame climate change as a political issue as financial resources exist for exploring research on climate change. Conversely, coverage of climate change news from poorer developing countries focus mainly on international relations. Thus, the effects of national macro economic variables that affects a country’s sociopolitical and economic development such as GDP reflects a country’s governance system in media coverage \cite{vu_what_2018}. Cultural theorists assert that, understanding such differential variations depicts how different entities interpret danger and respond to risk \cite{tansey_cultural_1999}. Inclusion of developing countries in scientific research provides an avenue for international support to such countries in responding to the effects of climate change.Yet studies attempting to explain cross-national variation in climate change public opinion is limited \cite{knight_public_2016}.

Public debates on the current consequences of climate change are discussed in both scientific and non-scientific mediums. 
% Jeremiah Osborne-Gowey: Need a better transition here between meteorological conditions and effect on mood (expressed as sentiment). Feels like we're missing a discussion (new paragraph?) just above here that discusses this. This might be where we plug in the text PAW is working on?
People can express their emotional states and feelings - sentiment - through physical, vocal and written expressions on mediums including newspapers, scientific articles, blogs and other online social media platforms. Twitter is one of the more popular social media platforms where people express their sentiments about any number of topics. Public posts on Twitter - called tweets - are widely available for public consumption and research, and offer a glimpse into collective social state. For example, text posted to these platforms can be used to gauge the public mood, assess opinions, measure brand affinity and for emergency planning and disaster response (cites). 
% Jeremiah Osborne-Gowey: Good to cite the top-cited public opinion papers, the Twitter flood study from CU, others from data science field including another Baylis paper
Sentiment is frequently employed as a correlate of emotional state or mood (cites). Weather events focus public attention in different ways of expression (cites). These data offer a unique opportunity to analyze how people express and respond to events in people's daily lives, including their exposure to ambient environmental conditions.

Here, we report on associations between meteorological conditions and expressed sentiment of public posts on Twitter from the United States of America (USA) between 2009 and 2019. This work builds on research from Baylis and others \cite{baylis_weather_2018} on weather impacts on expressed sentiment. In particular, we extend this work by including the temperature effects on sentiment in multiple regions across the USA while drawing in various climatic variables and socio-demographic data. In this research, we explore two primary questions. First, to what extent are ambient environmental conditions correlated with changes in expressed sentiment? Second, is this effect moderated by wealth?

Building on previous findings from Baylis and others, we hypothesize that local ambient environmental conditions have a strong connection how people are feeling as expressed in the sentiment of text messages posted on Twitter [H1]. Second, we hypothesize that the weather-sentiment link will be stronger in low-income areas. That is to say, people in lower income areas will be have lower sentiment scores than those in wealthier areas. [H2].

@@

\subsection{Effects by Time of Day}

In addition to providing data with high spatial resolution, twitter data also comes with very high temporal resolution as each tweet is time-stamped.  Thus, we were able to examine the impact of heat on sentiment for each hour of the day (See Fig. \ref{fig:ts-wbgt}).  We found that heat decreased expressed sentiment at nearly all hours of the day, and that heat had the largest effect on sentiment early in the morning, between 4am and 6am.  Heat had the weakest effects on sentiment at midday, and at 3am.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\linewidth]{../res/ts.png}
  \caption{Effects of rising temperatures on sentiment by hour of the day, with a 95\% confidence interval.  The value shown is the predicted change in VADER score for a 25-degree increase in temperatures.}
  \label{fig:ts-wbgt}
\end{figure}

@@

% Sentiment is a fuzzy indicator but we have BIG DATA
However, one limitation of data from twitter is that a tweet is not a precise measurement of a discrete mental health outcome and only provides a rough indicator a user's mental state based on the vocabulary in the tweet.  Moreover, while sentiment analysis algorithms have gotten increasingly sophisticated at estimating the mood in a body of text, a users expressed sentiment in a tweet is highly variable and is mostly affected by factors like current events and the user's personal life, with local weather conditions having only a small impact.  We overcome this limitation by using an extremely large data set of nearly 250 million tweets, allowing us to control for a wide variety of spatial and temporal effects to estimate population-level changes in sentiment in relation to the weather.

% Compare weekly changes in sentiment to weekly changes in suicides
We found that the change from optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to a heatwave of 25\textdegree C WBGT (36\textdegree C/97\textdegree F) is associated with an overall decrease in the VADER sentiment score of 0.01.  This is similar to the degree of change in sentiment over the course of a week from a Monday nadir (0.1268) to a Saturday peak (0.1372), a change in sentiment associated with a suicides increase of 23\% over the course of the study period \cite{CDC2021}.  Thus, while research on linkages between sentiment and other mental health outcomes like suicides and hospitalizations is still nascent, there are clear similarities in patters of sentiment and suicides at some time scales.  Moreover, both suicides and sentiment are affected by higher temperatures \cite{baylis_weather_2018, Burke2018Aug}.  Thus, while twitter sentiment is the only mental health indicator available at fine enough spatial and temporal scales to examine neighborhood and hourly effects, these changes in sentiment are representative of real human suffering, and are also very likely indicative of medical outcomes like suicides and hospitalizations, although this analysis cannot speak to those outcomes specifically.

@@

% Comment on Race, why no effect on hispanics?
We found that majority black neighborhoods are much more affected by heat than neighborhoods with a majority of any other race.  This is surprising, given that Hispanics can be marginalized in both income and housing.  This finding may be because the racial category of Hispanic is more broad and encompasses many more groups with more diverse histories and income levels than black Americans.  Additionally, there are many other marginalized groups in America, particularly indigenous people, that we did not have sufficient data to examine with respect to heat and mental health.

% Interpret Spatial Analysis
In our spatial analysis, we found that most cities exhibited both racial and income inequalities in the vulnerability of mental health to heat, with noticeable regional patterns in the strength of the effect.  Generally, the mid-west and mid-atlantic had the strongest inequalities, and the south-west had the weakest inequalities and an occasional inverted effect.  Because we used Wet Bulb Globe Temperature, our metric accounts for how differences in humidity and wind speed affect the human body's ability to cool itself through perspiration.  Thus, this regional differences in vulnerability are likely more due to cultural or infrastructural differences than to different environmental conditions.

@@

Where $y$ is the sentiment of a tweet, $t$ is the wet bulb globe temperature at the hour of the tweet, $p$ is a binary variable indicating whether it rained at the hour of the tweet, $s$ is the income shortwave radiation, or sunshine, in $W/m^2$, at the hour of the tweet, $\Phi$ is the spatio-temporal fixed effects, $\epsilon$ is the normally-distributed errors, and $f_t$ represented the segmented regression for $t$, with knots at -10\textdegree, 0\textdegree, 5\textdegree, 10\textdegree, 15\textdegree, 20\textdegree, and 25\textdegree C WBGT.

@@

Where $m$ is either the log-transformed average income in the census block where the tweet originated, or a dummy variable for the four racial categories.  We specify alternative models where income is a dummy variable in three bins, and were race is a continuous variable for the percent of a census block population that is what.  These alternative specifications yield similar results (See Supplement).  We also present the effects of rainfall and sunshine on sentiment by income and race in the Supplement.

>>>>

\begin{abstract}
%Limit 150 words: https://www.nature.com/nclimate/about/content
%Actually there are lots of examples of articles with > 150 words, but I think we should keep it concise
Higher temperatures associated with climate change are expected to have major impacts on human mental health.  Indeed, traditional analyses of heat and mental health outcomes using data collected by public agencies have found strong associations between elevated temperatures and outcomes such as suicides and mental-health related hospitalizations.  However, these studies, which typically use data reported at the city or county level, have found no difference in vulnerability based on income or race.  We overturn this finding and show that there are stark differences in vulnerability using a novel data source: expressed sentiment in 250 million geolocated tweets, matched with prevailing weather as well as neighborhood economic and demographic conditions.  We find that increased temperatures worsen expressed sentiment in all areas, but that this effects is much stronger in poor and black neighborhoods.

@@

\section{Main}
%Overview
Climate change is impacting many aspects of human well-being with these impacts projected to worsen in the coming years \cite{pachauri2014climate}.  A large body of research has assessed the ongoing impacts on climate change on outcomes like agriculture, economic growth, and physical health (NEED CITATIONS). Academic and medical researchers have highlighted the relative paucity of empirical studies into climate impacts on mental health at multiple scales and recently called for more \cite{Berry2018Apr, hayes_climate_2018}.  Researchers have begun to answer these calls, finding associations between deteriorating environmental conditions and a sense of "ecological grief" from their experienced loss \cite{Cunsolo2018Apr}, disasters and post-traumatic stress \cite{Waite2017Dec, Raker2019Dec}, as well as a variety of mental health impacts associated with warmer temperatures \cite{baylis_weather_2018, Mullins2019Dec} (PALINKAS and WONG 2020).  MAYBE A SENTENCE HERE HIGHLIGHTING THE GAP IN KNOWLEDGE ABOUT DISPARITIES IN MENTAL HEALTH IMPACTS, VULNERABLE POPULATIONS, issues of scale for signal to noise, etc.

%Heat and mental health
%https://www.pnas.org/content/115/43/10953
Higher temperatures are associated with many indicators of worsened mental health.  Multiple studies have found strong evidence that higher temperatures are associated with increases in suicides in the United States \citep{Burke2018Aug, Mullins2019Dec, Dixon2007May} with others demonstrating similar relationships in other locations around the world \citep{Qi2014Dec, Page2007Aug, Likhvar2011Jan}.  Higher temperatures have also been correlated with increases in hospitalization events related to both general mental health issues \cite{Obradovich2018Oct, Mullins2019Dec} and specific conditions like bipolar disorder and schizophrenia \citep{Lee2007Jan, Shapira2004Feb, Sung2013Feb, Gupta1992Jun, Hansen2008Oct}.

Temperature has been hypothesized to impact mental health through a number of pathways.  Work on biological mechanisms emphasize that there may be mental health effects from maintaining stable body temperature in high heat \cite{Lohmus2018Jul}.

Additionally, recent studies have found that nighttime temperatures affects sleep quality \citep{Obradovich2017May, Mullins2019Dec}.

% The pathways
Higher temperature associated with climate change could influence mental health directly by exposing people to trauma. It could also indirectly influence mental health by affecting (1) physical health and (2)community well-being \citep{RN1314}. For the direct influence pathway, extreme heat events or increasing temperatures have been associated with the increase in aggression,higher suicide rates and other hospital admissions \citep{RN1314,RN1316,RN1317,RN1318}. Heat exposures in working environments could also reduce people's capacity to deal with physical and mental task, and increase the risk of accidents, because of the excessive core body temperature and dehydration \citep{RN1319,RN1320,RN1321,RN1322}. The loss of work capacity would in turn result in loss of income, while the low income could also cause mental health problems \citep{Katz1997}.  As for the indirect pathway, first, because of the reciprocal causation relationship between physical health problems and mental health problems \citep{RN1323,RN1324} heat exposures associated with climate change along with other climate events and indirect health risks threatening physical health will directly influence mental health \citep{RN1325,RN1326}. Second, disordered temperature associated with climate change may destroy the economic in agricultural-production-dependent communities. For example, extreme heat reduces the work capacity of laborers in farm fields \citep{RN1320}, which further destroys agricultural-supported industries in local area\citep{RN1327}. The following economic pressures would undermine social capital and then lead to mental health problems.

Recent research conducted on the effects of higher temperatures on mental health conducted at large, multi-city scales has found surprisingly uniform vulnerability.  For example, focusing on suicides, Burke et al. found ``no significant difference in suicide response to temperature between rich and poor municipalities or counties" \cite{Burke2018Aug}.  Similarly, Mullins et al. find no effect of income as modifier of the effect of temperature on suicides, emergency department visits, or self-reported mental health status \cite{Mullins2019Dec}.  One major challenge for these studies is that they rely on data from public health agencies that is aggregated to the municipality or city level.  Thus, they must rely on between-county metrics of vulnerability and cannot take neighborhood effects into account.

%We would expect the poor to be more vulnerable, but this is not what we find
The poor and disadvantaged are much more exposed to undesirable temperatures.  They are more likely to work outside rather than indoors; they are more likely to rely on public transportation, bicycle, or walk instead of commute in their own air-conditioned car; and they are more likely to live in housing that is less insulated with poorer quality air-conditioning.  Thus, there are strong reasons to expect heterogeneities in the impact of heat on mental health outcomes, and these heterogeneities in vulnerability might be most pronounced at the neighborhood scale.  Indeed this is what we find for a variety of impacts of heat on physical health  \cite{Belanger2015Mar, Uejio2011Mar}.

While public health data on mental health outcomes is not available at the neighborhood scale, data from twitter can provide an important indicator of mental health at extremely precise spatial and temporal scales.  Many studies have used the mood expressed in tweets as an indicator of mental health \cite{Edo-Osagie2020Jul, Sinnenberg2016Dec}.  Such studies have used twitter data to identify the onset of Post Traumatic Stress Disorder in individuals, even before their formal diagnosis \cite{Reece2017Oct}, while similar methods with facebook data have been used to predict the onset of depression \cite{Eichstaedt2018Oct}.  Additionally, in London, day-to-day changes in mental health indicators derived from tweets have been associated with changes in mental health crisis episodes \cite{Kolliakou2020Feb}, while across the US more positive tweets are associated with a variety of metrics of human well being \cite{Mitchell2013May}.  Finally, the mood expressed in tweets has been strongly associated with local weather \cite{baylis_weather_2018}, and has been used as evidence to support the temperature-suicide relationship \cite{Burke2018Aug}.  Thus, we used a common sentiment metric to determine if neighborhood characteristics affect the vulnerability of mental health to higher temperatures.

@@

@@

% Sentiment is a fuzzy indicator but we have BIG DATA
However, one limitation of data from twitter is that a tweet is not a precise measurement of a discrete mental health outcome and only provides a rough indicator a user's mental state based on the vocabulary in the tweet.  Moreover, while sentiment analysis algorithms have gotten increasingly sophisticated at estimating the mood in a body of text, a user's expressed sentiment in a tweet is highly variable and is mostly affected by factors like current events and the user's personal life, with local weather conditions having only a small impact.  We overcome this limitation by using an extremely large data set of nearly 250 million tweets, allowing us to control for a wide variety of spatial and temporal effects to estimate population-level changes in sentiment in relation to the weather.

% Compare weekly changes in sentiment to weekly changes in suicides
We found that the change from optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to a heatwave of 25\textdegree C WBGT (36\textdegree C/97\textdegree F) is associated with an overall decrease in the VADER sentiment score of 0.01.  This is similar to the degree of change in sentiment over the course of a week from a Monday nadir (0.1268) to a Saturday peak (0.1372), a weekly change associated with a increase in suicides of 23\% over the course of the study period \cite{CDC2021}.  Thus, while research on linkages between sentiment and other mental health outcomes like suicides and hospitalizations is still nascent, there are clear similarities in patters of sentiment and suicides at some time scales.  Moreover, both suicides and sentiment are affected by higher temperatures \cite{baylis_weather_2018, Burke2018Aug}.  Thus, while twitter sentiment is the only mental health indicator available at fine enough spatial and temporal scales to examine neighborhood effects, these changes in sentiment are representative of real human suffering, and are also very likely indicative of medical outcomes like suicides and hospitalizations, although this analysis cannot speak to those outcomes specifically.

% Discuss the NOVELTY: quote Burke and Mullins, refute their interpretations, people have said XX, but we have shown ~XX.
% Variations in vulnerability also indicates ADAPTATION
Previous work conducted with county-scale data found no heterogeneities in mental health vulnerability by income.  Such analyses have led to conclusions that (1) the mental health effects of climate change will be uniform, least in developed countries, and (2) that adaptation with technologies like air-conditioning will not be enough.  For example, Mullins et al. conclude that ``high-temperature events will harm mental health everywhere" and that ``individuals have not been able to successfully reduce the negative effects of higher temperatures on mental health" \cite{Mullins2019Dec}.  Contrary to these analyses, we found large heterogeneities in vulnerability by neighborhood race and income.  This suggests that the mental health effects of climate change will not be uniform and, like other impact, will fall disproportionately on the poor and vulnerable.  More positively, it does suggest that adaptation is possible, and, for communities with infrastructure and working conditions similar those of high-income Americans, the effects of high temperatures on mental health can be largely mitigated.

@@

% Comment on Race, why no effect on hispanics?
We found that majority black neighborhoods are much more affected by heat than neighborhoods with a majority of any other race.  This is surprising, given that Hispanics can be marginalized in both income and housing.  This finding may be because the racial category of Hispanic is more broad and encompasses many more groups with more diverse histories and income levels than black Americans, or may also be due to that fact that more marginalized and heat-vulnerable Hispanic people were more likely to tweet in Spanish, which we excluded from our analysis.  Additionally, there are many other marginalized groups in America, particularly indigenous people, that we did not have sufficient data to examine with respect to heat and mental health.

% Interpret Spatial Analysis
In our spatial analysis, we found that heat affects sentiment in cities across the US, suggesting that higher baselines temperatures do not reduce vulnerability.   Additionally, most cities exhibited both racial and income inequalities in the vulnerability of mental health to heat, although the south-west had the weakest inequalities and an occasional inverted effect.  While this could suggest that heat inequality is related humidity, this is unlikely because we used Wet Bulb Globe Temperature, which accounts for how differences in humidity and wind speed affect the human body's ability to cool itself through perspiration.  Thus, this regional differences in vulnerability are likely more due to cultural or infrastructural differences than to different baseline environmental conditions.

@@

Where $y$ is the sentiment of a tweet, $t$ is the wet bulb globe temperature at the hour of the tweet, $p$ is a binary variable indicating whether it rained at the hour of the tweet, $s$ is the income shortwave radiation, or sunshine, in $W/m^2$, at the hour of the tweet, $\Phi$ is the spatio-temporal fixed effects, $\epsilon$ is the normally-distributed errors, and $f_t$ represented the segmented regression for $t$, with knots at -10\textdegree, 0\textdegree, 5\textdegree, 10\textdegree, 15\textdegree, 20\textdegree, and 25\textdegree C WBGT.  We estimate the 95\% confidence interval of our models using 80-fold bootstrapping.

@@

Where $m$ is either the log-transformed average income in the census block where the tweet originated, or a dummy variable for the four racial categories.  We specify alternative models where income is a dummy variable in three bins, and were race is a continuous variable for the percent of a census block population that is what.  These alternative specifications yield similar results.  We also examine the effects of rainfall and sunshine on sentiment by income and race.  These alternative specifications and analysis of rainfall and sunshine effects are available in the Supplement.

For our analyses by CSA, we subset the data in the model and run the model with the same fixed effects, although we fit a linear effect for $f_t()$.
